---
title: "Homogenous demand, conduct exercise"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Import & clean data, produce descriptive statistics

## Import data

First, import the data into a data frame, called coffee_raw:

```{r import}
coffee_raw <- read.csv("dutch_coffee.csv")

```

## Deflate prices using price index

Next, deflate the prices, save a new data frame, coffee:

```{r deflate, echo=FALSE}
library(dplyr)

coffee <- coffee_raw %>%
  mutate(cprice = cprice/oprice,
         tprice = tprice/oprice,
         bprice = bprice/oprice,
         wprice = wprice/oprice,
         incom = incom/oprice)

```

## Explore variations in price over time

Convert year/month variable into R time format, using as.Date and as.yearmon from the zoo package:

```{r date}
library(zoo)
coffee <- coffee %>%
  mutate(date = as.Date(as.yearmon(month)))
  
```


## Make some nice line graphs

Create simple line graphs of key price and income variables over time.

Basic plot of coffee price over time.

```{r cprice, echo=FALSE}
plot(cprice ~ date, coffee)


```

Basic plot of coffee bean price over time.

```{r bprice, echo=FALSE}
plot(bprice ~ date, coffee)

```
Basic plot of tea price against time:

```{r tprice, echo=FALSE}

plot(tprice ~ date, coffee)

```
Plot of wage prices over time:

```{r wprice}
plot(wprice ~ date, coffee)

```
Basic plot of income against time:

```{r income}
plot(incom ~ date, coffee)

```
Plot of other prices over time:

```{r oprice}
plot(oprice ~ date, coffee)

```

Now plot coffee retail price against the bean price using ggplot2 and geom_line:

```{r wholesale_retail}
library(ggplot2)
ggplot(coffee, aes(date)) +
  geom_line(aes(y = cprice), colour="red") + 
    geom_line(aes(y = bprice), colour="green") 


```
## Summary statistics

Produce a table of summary statistics for qu, cprice, tprice, wprice, bprice, incom using dplyr select:

```{r summary}

coffee %>%
  select(qu, cprice, tprice, wprice, bprice, incom) %>%
  summary()

```
Now consider summary statistics for quarter 1 alone:

```{r summary_q1}
coffee %>%
  filter(q1 == 1) %>%
  select(qu, cprice, tprice, wprice, bprice, incom) %>%
  summary()

```
Now do the same for q2, q3, and q4.

```{r summary_q2}
coffee %>%
  filter(q2 == 1) %>%
  select(qu, cprice, tprice, wprice, bprice, incom) %>%
  summary()

```

## Considering prices and quantities

First, use ggplot to produce a scatter graph of prices and quantities using geom_point:

```{r p_q, echo=FALSE}

coffee %>%
  ggplot(aes(x = qu)) +
    geom_point(aes(y = cprice))
  

```
Now do the same thing but only for the first quarter:

```{r filter}

coffee %>%
  filter(q1 == 1) %>%
  ggplot(aes(x = qu)) +
    geom_point(aes(y = cprice))

```

Now consider for q2, q3 and q4.

```{r filter1}

coffee %>%
  filter(q2 == 1) %>%
  ggplot(aes(x = qu)) +
    geom_point(aes(y = cprice))

```

```{r filter2}

coffee %>%
  filter(q3 == 1) %>%
  ggplot(aes(x = qu)) +
    geom_point(aes(y = cprice))

```

## Assess relationship between tea price and quantity demanded of coffee


```{r tprice_q}

coffee %>%
  ggplot(aes(x = qu)) +
    geom_point(aes(y = tprice))

```

## Price of coffee and labor costs

Consider the price of coffee and wages:

```{r cprice_w}
coffee %>%
  ggplot(aes(x = wprice)) + 
    geom_point(aes(y = cprice))

```

## Price of coffee and price of tea

Look at the relationship between the price of coffee and the price of tea:

```{r tprice_cprice}
coffee %>%
  ggplot(aes(x = tprice)) + 
    geom_point(aes(y = cprice))

```
```{r tprice_cprice}
coffee %>%
  filter(q4==1) %>%
  ggplot(aes(x = tprice)) + 
    geom_point(aes(y = cprice))

```

## Overall correlations

Produce a correlation matrix for qu, cprice, tprice, wprice, bprice, incom: 

```{r cor}
coffee %>%
  select(qu, cprice, tprice, wprice, bprice, incom) %>%
  cor()

```
## create log variables  

We need our variables to be in logs:

```{r logs, echo=FALSE, include = FALSE }
coffee <- coffee %>%
  mutate(ln_qu=log(qu),
         ln_cprice=log(cprice),
         ln_tprice=log(tprice),
         ln_bprice=log(bprice),
         ln_wprice=log(wprice),
         ln_incom=log(incom))
```

# Considering the basic relationships

## Basic demand regressions

Regress log quantity on log price and seasonal dummies and store as q_p1:

```{r q_p}
q_p1 <- lm(ln_qu ~ ln_cprice + q2 + q3 + q4 , coffee)
summary(q_p1)

```
Including tea price, save as q_p2:

```{r q_p2}
q_p2 <- lm(ln_qu ~ ln_cprice + q2 + q3 + q4 + tprice, coffee)
summary(q_p2)

```
Include income, save as q_p3:

```{r q_p3}
q_p3 <- lm(ln_qu ~ ln_cprice + q2 + q3 + q4 + tprice + incom, coffee)
summary(q_p3)

```
# Estimating theta

## Instrumental variable regressions

Run an instrumental variable regression, using prices, quantities, seasonal dummies and instrument, first without tea and income, and secondly with tea and income:

```{r iv1}
library("AER")

iv1 <- ivreg(ln_qu ~ q2 + q3 + q4 + ln_cprice | ln_bprice + q2 + q3 + q4 , data = coffee)
summary(iv1)

iv2 <- ivreg(ln_qu ~ q2 + q3 + q4 + incom + tprice + ln_cprice | ln_bprice + q2 + q3 + q4 + incom + tprice , data = coffee)
summary(iv2)

```
## Now lets show the adjusted learner index

Compute the adjusted lerner index using our cost information and our estimate of $\gamma$:

```{r lerner_adj, echo= FALSE}

k = 1.19 # kg of beans per kg of coffee
c0= 4 # by assumption (4 guilders)
gamma = -0.276

# obtain the cost function    
cost = c0 + k*coffee$bprice

# compute Lerner index
coffee$Ln= (coffee$cprice - cost) / coffee$cprice
  
# now compute adjusted Lerner index  
coffee$Ln_adj = -gamma * coffee$Ln 

# summary statistics for our adjusted Lerner index  
summary(coffee$Ln_adj)

# consider adjusted lerner for only q1, using filter
coffee %>%
  filter(q4 == 1) %>%
  select(Ln_adj) %>%
  summary(coffee$Ln_adj)


```
Now, consider for q2, q3, and q4.

## Estimating conduct parameter

First, regress prices on costs

```{r conduct}
conduct1 <- lm(cprice ~ cost + q1 + q2 + q3 -1, data = coffee) # omits intercept
summary(conduct1)

conduct2 <- lm(cprice ~ cost -1, data = coffee) # omits intercept
summary(conduct2)
```

From the FOC we have: $p=\frac{\gamma}{(\gamma+\theta)}c$
This is the estimated coefficient for cost.

```{r theta1}

b=1.47
theta=gamma*(1-b)/b
summary(theta)

n=1/theta
summary(n)    

```

## Linear demand 

$Q=\beta(\alpha-P)$

```{r}

linear1 <- ivreg(qu ~ cprice + incom + q1 + q2 + q3 | bprice + incom + q1 + q2 + q3, data = coffee)
summary(linear1)
linear2 <- ivreg(qu ~ cprice + q1 + q2 + q3 | bprice + q1 + q2 + q3, data = coffee)
summary(linear2)
```
Compute adjusted learner index:

```{r lerner_adj2}
coffee$gamma_l=-0.015*coffee$cprice/coffee$qu
coffee$Ln_adj_l=-coffee$gamma_l*coffee$Ln
```

Now summarise the lerner indices:

```{r l_adj1_summary}
coffee %>%
  select(Ln_adj_l) %>%
  summary()
```
Now summarise by quarters.

## Demand rotaters

Create seasonal demand rotators:

```{r rotators}
coffee <- coffee %>%
  mutate(cprice_q1=cprice*q1,
         bprice_q1=bprice*q1,
         cprice_q2=cprice*q2,
         bprice_q2=bprice*q2,
         cprice_q3=cprice*q3,
         bprice_q3=bprice*q3,
         cprice_q4=cprice*q4,
         bprice_q4=bprice*q4
  )

```

Can also use use demand rotators in regression:

```{r lm_rotators}
rotator <- ivreg(qu ~  q1 + q2 + q3 + (cprice + cprice_q1) | 
                   (bprice + bprice_q1) + q1 + q2 + q3 , data = coffee)
summary(rotator)

```




